---
title: "Practical Machine Learning - Course Project"
date: "Thursday, September 24, 2015"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

This is the solution to the course project in the practical machine learning class. The data is about persons doing barell lifts correctly or in 5 incorrect ways. We build a model to classify which of these 6 cases has been performed.

Training data needs to be downloaded from https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv and test data from https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv. We assume these files in the working directory and all required packages are installed.  


# First we read the training data

```{r}
trainingDF<-read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
```

# Cleaning the data
We remove attributes which are mainly NA, also the first 6 attributes which don't describe the quality of the exercise and therefor don't help the classification.
This cleaning needs to be repeaded with any new data that we want to classify later on.

```{r}
countNAs<-apply(is.na(trainingDF),2,sum)
cleanDF<-trainingDF[countNAs<10000]

cleanDF$X<-NULL
cleanDF$raw_timestamp_part_1<-NULL
cleanDF$raw_timestamp_part_2<-NULL
cleanDF$cvtd_timestamp<-NULL
cleanDF$new_window<-NULL
cleanDF$num_window<-NULL
```

# Building a random forest model
We load the caret library, set a seed for reproduceability and split the data in a training and test set.

```{r}
library(caret)
set.seed(123456789)
inTrain = createDataPartition(cleanDF$classe, p = 3/4)[[1]]
training = cleanDF[inTrain,]
testing = cleanDF[-inTrain,]
modelFit <- train(classe ~.,data= training, method="rf")
```

# prediction on the test data set and check confusion matrix

```{r}
predictions<-predict(modelFit,newdata = testing)
confusionMatrix(predictions,testing$classe)
```
Therefore the accuracy is 99%.

# CLassify the 20 observations in the test data
First we read the data and clean it as before.

```{r}
testingDF<-read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
testingDF<-testingDF[countNAs<10000]

testingDF$X<-NULL
testingDF$raw_timestamp_part_1<-NULL
testingDF$raw_timestamp_part_2<-NULL
testingDF$cvtd_timestamp<-NULL
testingDF$new_window<-NULL
testingDF$num_window<-NULL
```

The we use the model to classify:

```{r}
classification<-predict(modelFit,newdata = testingDF)
```

Now we export the results as suggested in the project description.

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(classification)
```